---
title: "cite-bike-analysis-4"
author: "Amelia Chu"
date: "4/1/2018"
output: html_document
---
Let's import the library `forecast` as suggested by Professor Perry.

```{r}
library("forecast")
```
Let's read in the dataset. The orginal data was pulled from the Citi Bike website() by downloading the historical daily data in quarterly reports.

This data was then aggregated into monthly data using numpy and pandas dataframes. This is the csv export.
```{r}
data <- read.csv("/Users/chuamelia/Google Drive/unemp.csv")
date <- as.Date(data$date)
time <- 1:length(date)
trips <- data$unemp

log.trips <- log(trips)
#diff.log.trips <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, diff(log.trips, lag = 12))
#diff2.log.trips <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, diff(diff.log.trips, lag =12))
```

Plotting four timeseries to help inform decision to log or difference the data.
```{r}
plot(time, data$Trips, type="l", xlab="Date", ylab="Trips")
plot(date, log.trips, type="l", xlab="Date", ylab=" Log Trips")
#Calculating differences at lag 12 appears to "get rid" of the seasonal pattern. There does not seem to be must difference betweenn differenceing once and twice.
plot(date, diff.log.trips, type="l",
     xlab="Date", ylab="Diff Log Trips")
plot(date, diff2.log.trips, type="l",
     xlab="Date", ylab="Diff 2 Log Trips")
```

```{r}
plot(date, diff2.log.trips, type="l", col="blue"
  )
lines(date, diff3.log.trips, type="l",col="green"
    )
```

```{r}
acf(log.trips, na.action = na.pass) #dies down, returns at lag 10
pacf(log.trips, na.action = na.pass) #hangs at 1
```
```{r}
# ACF and PACF
acf(diff.log.trips, na.action = na.pass) # dies down
pacf(diff.log.trips, na.action = na.pass) # cuts off at 1 or 3 might be worth trying AR(1) or AR(2)
```
```{r}
# ACF and PACF
acf(diff2.log.trips, na.action = na.pass) # dies down
pacf(diff2.log.trips, na.action = na.pass) # cuts off at 1 or 3 might be worth trying AR(1) or AR(2)
```
```{r}
# Without constant:
fit.00 <- Arima(log.trips, c(0, 0, 0), include.constant=FALSE)
fit.01 <- Arima(log.trips, c(0, 0, 1), include.constant=FALSE)
fit.02 <- Arima(log.trips, c(0, 0, 2), include.constant=FALSE)
fit.10 <- Arima(log.trips, c(1, 0, 0), include.constant=FALSE)
#fit.11 <- Arima(log.trips, c(1, 0, 1), include.constant=FALSE)
fit.12 <- Arima(log.trips, c(1, 0, 2), include.constant=FALSE)
fit.20 <- Arima(log.trips, c(2, 0, 0), include.constant=FALSE)
fit.21 <- Arima(log.trips, c(2, 0, 1), include.constant=FALSE)
fit.22 <- Arima(log.trips, c(2, 0, 2), include.constant=FALSE)

# With constant:
fit.00c <- Arima(log.trips, c(0, 0, 0), include.constant=TRUE)
fit.01c <- Arima(log.trips, c(0, 0, 1), include.constant=TRUE)
fit.02c <- Arima(log.trips, c(0, 0, 2), include.constant=TRUE)
fit.10c <- Arima(log.trips, c(1, 0, 0), include.constant=TRUE)
fit.11c <- Arima(log.trips, c(1, 0, 1), include.constant=TRUE)
fit.12c <- Arima(log.trips, c(1, 0, 2), include.constant=TRUE)
fit.20c <- Arima(log.trips, c(2, 0, 0), include.constant=TRUE)
fit.21c <- Arima(log.trips, c(2, 0, 1), include.constant=TRUE)
fit.22c <- Arima(log.trips, c(2, 0, 2), include.constant=TRUE)

# Summarize Results
models <- data.frame(p = rep(c(0, 0, 0, 1, 1, 1, 2, 2, 2), 2),
                     d = rep(1, 18),
                     q = rep(c(0, 1, 2), 6),
                     include.constant = c(rep(FALSE, 9), rep(TRUE, 9)),
                     loglik = c(fit.00$loglik, fit.01$loglik, fit.02$loglik,
                                fit.10$loglik, 0, fit.12$loglik,
                                fit.20$loglik, fit.21$loglik, fit.22$loglik,
                                fit.00c$loglik, fit.01c$loglik, fit.02c$loglik,
                                fit.10c$loglik, fit.11c$loglik, fit.12c$loglik,
                                fit.20c$loglik, fit.21c$loglik, fit.22c$loglik),
                     aicc = c(fit.00$aicc, fit.01$aicc, fit.02$aicc,
                                fit.10$aicc,0, fit.12$aicc,
                                fit.20$aicc, fit.21$aicc, fit.22$aicc,
                                fit.00c$aicc, fit.01c$aicc, fit.02c$aicc,
                                fit.10c$aicc, fit.11c$aicc, fit.12c$aicc,
                                fit.20c$aicc, fit.21c$aicc, fit.22c$aicc)
                     )
print(models, digits=6)
```

```{r}
fit <- Arima(diff.log.trips, c(1, 0, 0), include.constant=TRUE)
print(fit)
```

```{r}
# Hint: use Box.test function.  You must call this function four times,
# for lags 12, 24, 36, and 48.
Box.test(diff.log.trips, lag = 12, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
Box.test(diff.log.trips, lag = 24, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
Box.test(diff.log.trips, lag = 36, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
Box.test(diff.log.trips, lag = 48, type = c("Box-Pierce", "Ljung-Box"), fitdf = 0)
```

```{r}
# Add TS plot, ACF, and PACF
resid.log.trips <- residuals(fit)
# Time series plot
plot(as.Date(data$MonthYear), resid.log.trips, type="l",
     xlab="Date", ylab="Log Trips Residuals")
# ACF and PACF
acf(resid.log.trips, na.action = na.pass)
pacf(resid.log.trips, na.action = na.pass)
```
```{r}
plot(forecast(fit, h=50, level=95), col=2 )

```
